---
title: "ADA Homework 3"
output: html_document
---

**Challenge 1**

```{r}
library(curl)
library(readr)
library(broom)
library(ggplot2)
library(ggpubr)
library(dplyr)
f <- curl("https://raw.githubusercontent.com/raneyej/ada-2021-datasets/main/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",")
head(d)
```

Generate a linear model: 

```{r}
bs_lm <- lm(WeaningAge_d ~ Brain_Size_Species_Mean, data = d)
summary(bs_lm)

bs_lm_log <- lm(log(WeaningAge_d) ~ log(Brain_Size_Species_Mean), data = d)
summary(bs_lm_log)
```

Generate scatterplots for the models: 

```{r}
brain_plot <- ggplot(data = d, aes(x = Brain_Size_Species_Mean, y = WeaningAge_d)) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, color = "blue") +
  stat_regline_equation(label.y = 1500, aes(label = ..eq.label..)) +
  ggtitle(label = "Brain size and weaning age") + xlab("Brain size") + ylab("Weaning age in days")
brain_plot

log_brain_plot <- ggplot(data = d, aes(x = log(Brain_Size_Species_Mean), y = log(WeaningAge_d))) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, color = "blue") +
  stat_regline_equation(label.y = 8, aes(label = ..eq.label..)) +
  ggtitle(label = "Log transformed brain size and weaning age") + xlab("Log brain size") + ylab("Log weaning age in days")
log_brain_plot
```

Coefficients: 

```{r}
tidy(bs_lm)
tidy(bs_lm_log)

beta0 <- tidy(bs_lm) %>%
  filter(term == "(Intercept)") %>%
  pull(estimate)
beta0
beta1 <- tidy(bs_lm_log) %>%
  filter(term == "log(Brain_Size_Species_Mean)") %>%
  pull(estimate)
beta1
```

Hypothesis test: 

```{r}
beta1 == 0
beta1!=0
```

The slope estimate for the regular model is 2.64, and the slope estimate for the log-transformed model is 0.57. Both p-values are much less than 0.001 (1.59e-26 and 1.99e-35, respectively), so we can reject the null hypothesis and assume that there is a relationship between brain size and weaning age. 

Confidence intervals: 

```{r}
ci_beta1 <- predict(bs_lm_log, newdata = data.frame(Brain_Size_Species_Mean = 2.64), interval = "confidence", level = 0.90)
ci_beta1

ci_bs_lm <- predict(bs_lm, newdata = data.frame(Brain_Size_Species_Mean <- d$Brain_Size_Species_Mean), interval = "confidence", level = 0.90)
ci_bs_lm

ci_bs_lm_log <- predict(bs_lm_log, newdata = data.frame(Brain_Size_Species_Mean <- d$Brain_Size_Species_Mean), interval = "confidence", level = 0.90)
ci_bs_lm_log

ci_array <- cbind(d$Brain_Size_Species_Mean, ci_bs_lm)
ci_df <- as.data.frame(ci_array)
ci_df

x1 <- d$Brain_Size_Species_Mean
x1
x2 <- log(x1)
x2

ci_array2 <- cbind(x2, ci_bs_lm_log)
ci_df2 <- as.data.frame(ci_array2)
ci_df2

ci_plot1 <- ggplot(data = d, aes(x = Brain_Size_Species_Mean, y = WeaningAge_d)) +
  geom_point() +
  geom_line(data = ci_df, aes( x = V1, y = fit, color = "Fit")) +
  geom_line(data = ci_df, aes(x = V1, y = lwr, color = "Lower")) +
  geom_line(data = ci_df, aes(x = V1, y = upr, color = "Upper")) +
  stat_regline_equation(label.y = 1500, aes(label = ..eq.label..)) +
  ggtitle(label = "Brain size and weaning age") + xlab("Brain size") + ylab("Weaning age in days")
ci_plot1

ci_plot2 <- ggplot(data = d, aes(x = log(Brain_Size_Species_Mean), y = log(WeaningAge_d))) +
  geom_point() +
  geom_line(data = ci_df2, aes( x = x2, y = fit, color = "Fit")) +
  geom_line(data = ci_df2, aes(x = x2, y = lwr, color = "Lower")) +
  geom_line(data = ci_df2, aes(x = x2, y = upr, color = "Upper")) +
  stat_regline_equation(label.y = 7, aes(label = ..eq.label..)) +
  ggtitle(label = "Brain size and weaning age") + xlab("Log brain size") + ylab("Log weaning age in days")
ci_plot2
```

Predict weaning age for a brain size of 750g: 

```{r}
prediction <- predict(bs_lm, newdata = data.frame(Brain_Size_Species_Mean = 750), interval = "prediction", level = 0.90)
prediction
```

The predicted age of weaning for a primate with a brain size of 750 grams is 2,110 days, or about 6 years. I am not confident in this estimate, because although their is a relationship between brain size and weaning age, 750 grams is much larger than the largest average brain size in the data set (491 grams). Because it is so much larger, it is possible that the model might not predict weaning age accurately. 

I think that the transformed model is better, because the relationship between brain size and weaning age is much more obvious than in the regular model. 

*Challenge 2* 
```{r}
gs_lm_log <-  lm(log(MeanGroupSize) ~ log(Body_mass_female_mean), data = d)
gs_lm_log
summary(gs_lm_log)
tidy(gs_lm_log)
```

The slope is 0.506, and the y-intercept is -1.78. 

```{r}
gs_beta0 <- tidy(gs_lm_log) %>%
  filter(term == "(Intercept)") %>%
  pull(estimate)
gs_beta0
gs_beta1 <- tidy(gs_lm_log) %>%
  filter(term == "log(Body_mass_female_mean)") %>%
  pull(estimate)
gs_beta1


boot_betas <- data.frame(gs_beta0_boot = 1:1000, gs_beta1_boot = 1:1000)
n = nrow(d)
for(i in 1:1000) {
  s <- sample_n(d, size = n, replace = TRUE)
  gs_boot <- lm(log(MeanGroupSize) ~ log(Body_mass_female_mean), data = s)
  betas <- gs_boot$coefficients
  gs_beta0_boot <- as.numeric(betas[1])
  gs_beta1_boot <- as.numeric(betas[2])
  boot_betas$gs_beta0_boot[[i]] <- gs_beta0_boot
  boot_betas$gs_beta1_boot[[i]] <- gs_beta1_boot
}
boot_betas
```

Histograms of distributions: 

```{r}
hist_beta0 <- hist(boot_betas$gs_beta0_boot, main = "Distribution of intercepts", xlab = "β0")
hist_beta1 <- hist(boot_betas$gs_beta1_boot, main = "Distribution of slopes", xlab = "β1")
hist_beta0
hist_beta1
```

Standard errors: 

```{r}
beta0_se <- sd(boot_betas$gs_beta0_boot)
beta0_se
beta1_se <- sd(boot_betas$gs_beta1_boot)
beta1_se
```

Confidence intervals: 

```{r}
quantile(boot_betas$gs_beta0_boot, prob = 0.025)
quantile(boot_betas$gs_beta0_boot, prob = 0.975)

quantile(boot_betas$gs_beta1_boot, prob = 0.025)
quantile(boot_betas$gs_beta1_boot, prob = 0.975)
```

Comparison of standard errors: 

```{r}
se_comp <- data.frame(Method = c("b0_lm", "b0_se", "b1_lm", "b1_se"), 
           Value = c(0.438, 0.483, 0.0556, 0.0616))
se_comp
```

Overall, the standard errors generated by bootstrapping are similar to the standard errors generated with the lm() function.

Comparison of confidence intervals: 

```{r}
cis_comp <- data.frame(Method = c("qt_b0_lower", "lm_b0_lower", "qt_b0_upper", "lm_b0_upper", "qt_b1_lower", "lm_b1_lower", "qt_b1_upper", "lm_b1_upper"), 
                       Value = c(-2.684, -2.642, -0.837, -0.912, 0.378, 0.396, 0.616, 0.625))
cis_comp
```

The confidence intervals generated by the lm() function are also very similar to the CIs generated using the quantile method. 

*Challenge 3*
 
```{r}
boot_lm <- function(d, model, conf.level = 0.95, reps = 1000) {
  df <- data.frame(ce = c("Beta0", "Beta1"), ce_value = c(0,0), ce_se = c(0,0), ci_upper = c(0,0), ci_lower = c(0,0), beta_bs_mean = c(0,0), beta_bs_se = c(0,0), bs_ci_upper = c(0,0), cs_ci_lower = c(0,0))
  mod <- lm(eval(parse(text = model)), data = d)
  mod_tid <- tidy(mod)
  df$ce[1] <- as.numeric(mod_tid[1,2])
  df$ce[2] <- as.numeric(mod_tid[2,2])
  df$ce_value[1] <- as.numeric(mod_tid[1,3])
  df$ce_value[2] <- as.numeric(mod_tid[2,3])
  mod_ci <- confit(model, level = conf.level)
  df$ci_upper[1] <- as.numeric(mod_ci[1,1])
  df$ci_upper[2] <- as.numeric(mod_ci[1,2])
  df$ci_lower[1] <- as.numeric(mod_ci[2,1])
  df$ci_lower[2] <- as.numeric(mod_ci[2,2])
  set.seed(1)
  bootstrap <- data.frame(beta0 = 1:reps, beta1 = 1:reps)
  n = nrow(d)
  for(i in 1:reps){
    s <- sample_n(d, size = n, replace = TRUE)
    mod_boot <- lm(eval(parse(text = model)), data = s)
    ce_boot <- mod_boot$coefficients
    beta0 <- as.numeric(ce_boot[1])
    beta1 <- as.numeric(ve_boot[2])
    mod_boot$beta0[[i]] <- beta0
    mod_boot$beta1[[i]] <- beta1
  }
  df$beta_bs_mean[1] <- mean(mod_boot$beta0)
  df$beta_bs_mean[2] <- mean(mod_boot$beta1)
  df$beta_se_mean[1] <- sd(mod_boot$beta0)
  df$beta_se_mean[2] <- sd(mod_boot$beta1)
  alpha <- 1 - conf.level
  df$bs_ci_upper[1] <- quantile(mod_boot$beta0, 1 - (alpha/2))
  df$bs_ci_upper[2] <- quantile(mod_boot$beta1, 1 - (alpha/2))
  df$bs_ci_lower[1] <- quantile(mod_boot$beta0, (alpha/2))
  df$bs_ci_lower[2] <- quantile(mod_boot$beta1, (alpha/2))
  df
}
```

Models: 

```{r, error = TRUE}
group_size <- boot_lm(d, "log(MeanGroupSize)" ~ "log(Body_mass_female_mean)")
day_length <- boot_lm(d, "log(DayLength_km)" ~"log(Body_mass_female_mean)")
day_length2 <- boot_lm(d, "log(DayLength_km" ~ "log(Body_mass_female_mean)" + "log(MeanGroupSize)")
```